An Implementation of Reference Counting for a Python to x86 Compiler

### 2011-12-15 paper/paper.tex
fix some grammar

The performance characteristics of each strategy can vary considerably, depending on implementation.   Tracing garbage collectors can sometimes be associated with long pause times, where application work must be temporarily stopped to allow the entire contents of the heap to be scanned by the garbage collector.  These long pause times have largely been solved with new algorithms that improve upon the original naive mark-and-sweep algorithm.  It is typical that these algorithms trade-off throughput for responsiveness.  However, these pause times can still be an issue, especially in real time systems, where a hard guarantees on latency are needed.  Reference counting has been discussed as a solution to this problem since it incrementally deallocates objects at the program point where all references are destroyed, but classic implementations of reference counting can have serious performance concerns when one deallocation results in cascading deallocations\cite{boehm}.

@@

\subsubsection{Decrementing reference counts}
An assignment to a variable should result in a decrement of the reference count for the object referred to by the variable before the assignment.  However, this can't be done when the variable doesn't refer to anything, which is true prior to the initial assignment.  In fact, trying to even access such a variable before it is assigned to can result in undefined results, since there is no guarantee on the contents of the variable (it could be garbage or a value from a previous stack frame).   To prevent a call to \afunction{dec\_ref\_ctr} on a variable before the initial assignment, we initially tried to statically determine when the first assignment occurred.  This approach failed when we realized that for some programs, it is not possible to determine where the initial assignment occurs at compile-time.  Listing~\ref{lst:initialassign} shows an example program where the initial assignment to the variable \avariable{x} cannot be determined statically.  If the input provided by the user evaluates to \aliteral{True}, then the initial assignment to $x$ occurs at line 2.  However, if the input evaluates to \aliteral{False}, then the initial assignment to \avariable{x} occurs at line 3.  If we transform the input program so that a call is made to decrement the reference count for \avariable{x} immediately prior to the assignment on line 3, then there is the possibility that this will cause the program to crash if the user provides an input that evaluates to \aliteral{False}.

>>>>

The performance characteristics of each strategy can vary considerably, depending on implementation.   Tracing garbage collectors can sometimes be associated with long pause times, where application work must be temporarily stopped to allow the entire contents of the heap to be scanned by the garbage collector.  These long pause times have largely been solved with new algorithms that improve upon the original naive mark-and-sweep algorithm.  It is typical that these algorithms trade-off throughput for responsiveness.  However, these pause times can still be an issue, especially in real time systems, where hard guarantees on latency are needed.  Reference counting has been discussed as a solution to this problem since it incrementally deallocates objects at the program point where all references are destroyed, but classic implementations of reference counting can have serious performance concerns when one deallocation results in cascading deallocations\cite{boehm}.

@@

\subsubsection{Decrementing reference counts}
An assignment to a variable should result in a decrement of the reference count for the object referred to by the variable before the assignment.  However, this can't be done when the variable doesn't refer to anything, which is true prior to the initial assignment.  In fact, trying to even access such a variable before it is assigned produces undefined behavior since there is no guarantee on the contents of the variable (it could be garbage or a value from a previous stack frame).   To prevent a call to \afunction{dec\_ref\_ctr} on a variable before the initial assignment, we initially tried to statically determine when the first assignment occurred.  This approach failed when we realized that for some programs, it is not possible to determine where the initial assignment occurs at compile-time.  Listing~\ref{lst:initialassign} shows an example program where the initial assignment to the variable \avariable{x} cannot be determined statically.  If the input provided by the user evaluates to \aliteral{True}, then the initial assignment to $x$ occurs at line 2.  However, if the input evaluates to \aliteral{False}, then the initial assignment to \avariable{x} occurs at line 3.  If we transform the input program so that a call is made to decrement the reference count for \avariable{x} immediately prior to the assignment on line 3, then there is the possibility that this will cause the program to crash if the user provides an input that evaluates to \aliteral{False}.
